---
title: "PROJECT"
author: "Anastasia Buianova"
output:
  html_document: default
  pdf_document: default
---
# "Distribution of Number in Russian Constructions with names of Body Parts"
### HSE, Moscow. Course: Data Analysis and Visualisation.

### Content:

  1. Introduction;  
  2. Data Set;  
  3. Research Hypothesis;  
  4. Descriptive Statistics and Visualization;  
  5. Multi-factor analysis;  
  6. Linguistic interpretation;  
  7. References.
  
## 1. Introduction.

This project is made in a frame of my Bachelor and Master research "Constructions with names of body part from the typological perspective". The main aim of the research is to determine how the body parts are categorized through the prism of constructions. It is expected that the division of the human body into parts represented in constructions does not coincide with the usual concepts that can be obtained from a picture.

The current stage of the research is based only on the Semantic analysis of a whole construction. However, I would like to have a look at some grammatical and morphological features of the constructions. This distribution seems to have a linguistic value as well.

All the Russian constructions in the research are collected in the Russian National Corpus (НКРЯ, ruscorpora.ru) as 2-grams or 3-grams.

## 2. Dataset.

The dataset was especially designed for this project but with an intention to use it later as a part of my research. That's why it contains more information than needed for the project. It has 400 constractions for upper and lower limbs and following 13 columns (some of them are not relevant as it mentioned, so we can ignore them):

LANG - a language of a construction (Russian, Czech, German or English, in our case only Russian);
ZONE - a zone of a body part - upper limb (arm) or lower limb (leg);
BP_NAME - a body part name used in a construction;
SUBJ/OBJ - showes if a body part is a subject or an object in a construction; 
CONSTRUCTIONS - list of constructions;
GROUP - a construction group (IDIOM, A, B, or C - according to the research hypothises it makes sense for body part conceptualization);
VERB - a verb used in a construction;
PF/IMPF - showes if a verb is perfective or imperfective;
CLASS - class of a construction, based on Semantics of verbs and prepositions (52 classes);
CLUSTER - a class of construction classes (5 classes: MOVEM, PLACE, SENSE, IDIOM, OTHER);
NUM - number of a body part name might be used in a construction (according to the corpus data);
DO/IO - literally showes if a construction is a 2-gramm or 3-gramm (a body part is a direct object or an indirect object);
PREP - a preposition of a construction, if it's a 3-gramm;
CASE - a case of a body part.

All these features are needed to make a detailed grammatical and morphological analysis of body parts constructions in different languages, that's why they are in the dataset. As I mentioned above, some of them are irrelevant for this project (such as LANG, BP_NAME, SUBJ/OBJ, VERB, PF/IMPF, DO/IO, PREP, CASE).

Two varibles CLASS and CLUSTER have partly the same funtion. They are semantic concepts of [verb + preposition] part of a construction. The first one has a huge varience that makes it difficult for the statistical analysis of such a small dataset. And the second one clusterizes and unites these 52 verb classes into bigger concepts:

MOVEM: BARE, BARE+ORIENT, BEND, CLOSE, DOWN, FALL, FIX, FREE, GLUE, GO, MOVEM, OPEN, OUT, PAINT, RASP, REMOVE, SET, SHAKE, SHOW, SIT, TAKE, THROUGH, TURN, UP, WRAP - "Movements"
IDIOM: IDIOM - Idiomatic constructions
PLACE: CARRY, LAY, LAY+ORIENT, LAY+STATE, LEAN, LOC, ORIENT, PLACE - "Location"
SENSE: BEAT, EMOT, HURT, SENSE, STATE, STAY, TASTE, TOUCH - "Senses, states, and emotions"
OTHER: COSM, DIRTY, DRESS, DRESS DOWN, NO CLASS, SPIT, WASH - Other classes that can not be interpreted as movements, senses or location


##3. Research Hypothesis.

The project hypothesis is that the distribution of body part number (only Sg., only Pl., or both Sg./Pl.) depends on a verb cluster and zone - lower limbs tend to be used more in plural rather than upper limbs.

## 4. Descriptive Statistics, Statistical Significance and Visualization.

### Libraries.

```{r include = TRUE}
library(tidyverse)
library(vcd)
library(ggrepel)
library(nnet) #Multinominal Logistic Regression
library(FactoMineR) #MCA
library(factoextra) #MCA
```

### Step 1. Norminalizing the dataset.

That's how the dataset does look. However, as you remember, we don't need a half of the variables.

```{r include = TRUE}
#In the case if unicode is broken:
#Sys.setlocale(category = 'LC_ALL', 'en_US.UTF-8')

bp_original <- read_csv('/Users/NP/Desktop/PROJECT/DATASET.csv')

#In the case if the first variant doesn't work:
#bp <- read.csv('/Users/NP/Desktop/DATASET.csv', sep=',')

head(bp_original)
```

So that's what we actually need now.

```{r include = TRUE}
bp <- bp_original %>%
  select(2, 5:6, 9:11)

head(bp)
```

### Step 2. Descriptive statistics

Here we do look at the general statistics of the data. 
Table 1 showes the statistical numbers of the grammatical number distribution. There are  
  * 142 constructions which might have only plural number (NB: according to the corpus) such as "сшибить с ног", *"сшибить с ноги";  
  * 104 only for singular number such as "задеть плечом", *"задеть плечами";  
  * and 154 where it's possible to use both numbers - singular and plural, such as "махать рукой"/"махать руками".  
  
Table 2 showes how many constructions each of the verb classes has.

Table 3: the same thing as two previous ones, but for verb clusters. There are:  
  * 161 constructions with the meaning of movement - "MOVEM" cluster;			
  * 108 for senses, states and emotions - "SENSE";			
  * 57 for locations - "PLACE";			
  * 47 idiomatic constructions - "IDIOM";			
  * and 27 are the cluster of "OTHER"."	

Tables 4, 5, and 6 show the statistical distribution of the cluster, class, zone and grammatical number in different combinations.

```{r include = TRUE}
#Table 1.
bp %>%
  group_by(NUM) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

#Table 2.
bp %>%
  group_by(CLASS) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

#Table 3.
bp %>%
  group_by(CLUSTER) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

#Table 4.
bp %>%
  group_by(CLUSTER, CLASS, NUM) %>%
  summarise(count=n()) %>%
  arrange(NUM, desc(count))

#Table 5.
bp %>%
  group_by(CLUSTER, NUM) %>%
  summarise(count=n()) %>%
  arrange(NUM, desc(count))

#Table 6.
bp %>%
  group_by(CLUSTER, ZONE, NUM) %>%
  summarise(count=n()) %>%
  arrange(NUM, desc(count))
```

### Step 3. Descriptive statistics visualization.

The graph 1 showes the correlation between a verb cluster and a grammatical number of a used in a construction body part.

```{r include = TRUE}
bp %>%
  group_by(CLUSTER, NUM) %>%
  summarise(count=n()) %>%
  ggplot(aes(CLUSTER, NUM, label = count, size = count)) +
  geom_point(aes(size = count), color = "lightblue") + 
  geom_text(size = 4) +
  guides(size = F, color = F, fill = F) +
  scale_size(range = c(10, 30)) +
  labs(title = "Correlation between a verb cluster and number of a body part") +
  xlab("CLUSTER") +
  ylab("NUM")

```

The graph 2 showes the correlation between a verb cluster and a grammatical number of a used in a construction body part according to the body zone - lower limbs or upper limbs.
```{r include = TRUE}
bp %>%
  group_by(CLUSTER, ZONE, NUM) %>%
  summarise(count=n()) %>%
  ggplot(aes(CLUSTER, NUM, label = count, size = count)) +
  geom_point(aes(size = count), color = "lightblue") + 
  geom_text(size = 4) +
  guides(size = F, color = F, fill = F) +
  scale_size(range = c(10, 30)) +
  facet_wrap(~ZONE) +
  guides(fill = F) +
  labs(title = "Correlation between a verb cluster and number of a body part according to the body zone") +
  xlab("CLUSTER") +
  ylab("NUM")
```


### Step 4. Statistical significance, Pearson's Chi-squared Test

On this stage I would like to have a look, if a data I have is actually statistical significant. The main hypothesis of this project is that the distribution of a body part number (Sg., Pl., or Sg./Pl.) depends on a verb cluster and a zone. That's why I would like to check these two associations CLUSTER ~ NUM, ZONE ~ NUM, and two more to have a look if two previous ones are not the only dependant associations: CLASS ~ NUM, GROUP ~ NUM. I'm goint to use Pearson's Chi-squared Test.

H0: `CLUSTER` ~ `NUM`,  `ZONE` ~ `NUM` are independant
H1: `CLUSTER` ~ `NUM`,  `ZONE` ~ `NUM` are related.

Same for `CLASS` ~ `NUM`, `GROUP` ~ `NUM`.

```{r include = TRUE}
#Doesn't seem to be a statistocal significant association due to the small cells value
chisq.test(bp$CLASS, bp$NUM)

chisq.test(bp$CLUSTER, bp$NUM)

chisq.test(bp$ZONE, bp$NUM)
#fisher.test(bp$ZONE, bp$NUM)

#Doesn't seem to be a statistocal significant association due to the small cells value
chisq.test(bp$GROUP, bp$NUM)
```

CLASS ~ NUM, GROUP ~ NUM, to these associations Pearson's Chi-squared Test can not be applied due to the small cell values in the contingency table (because it can be incorrect). The only trusted pairs are the ones we actually look at CLUSTER ~ NUM, ZONE ~ NUM, for which p-Value is less than 5%. So we reject the null hypothesis and conclude that `CLUSTER`, `ZONE` and `NUM` have a significant relationship (dependant, the data is proved as a statistically significant associations).




## 5. Multi-factor analysis. 

### Multinominal Logistic Regression. 

Here the multinominal logistic regression model is used to predict ClUSTER by NUM (and later ZONE by NUM). It works like series of logistic regressions, comparing every two levels of an dependant varible.

The first association CLUSTER ~ NUM is not well-predictable (probably due to the lack of the data, for instance for the "OTHER"-cluster the coefficient is -1,4 and it has only 27 contructions from 400).
```{r include = TRUE}
#For CLUSTER ~ NUM association
cluster_mlg <- multinom(bp$CLUSTER ~ bp$NUM)
summary(cluster_mlg)
```

```{r include = TRUE}
#For ZONE ~ NUM association
zone_mlg <- multinom(bp$ZONE ~ bp$NUM)
summary(zone_mlg)
```

### Multiple correspondence analysis (MCA).

There are no good associations between some of the variable categories such as MOVEM, OTHER, and SENSE, as well as for LOWER LIMB.

```{r include = TRUE}
cluster_mca <- bp[, 5:6]
clust_res.mca <- MCA(cluster_mca, graph = FALSE)
clust_res.mca
clust_res.mca$var
```


```{r }
fviz_mca_var(clust_res.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, # Avoid text overlapping (slow)
             ggtheme = theme_minimal())

fviz_mca_var(clust_res.mca, alpha.var="contrib",
             repel = TRUE,
             ggtheme = theme_minimal())
```

```{r include = TRUE}
zone_mca <- bp[, c(1,6)]
zone_res.mca <- MCA(zone_mca, graph = FALSE)
zone_res.mca
zone_res.mca$var
```

##  6. Linguistic interpretation.  

The distrubution of grammatical number of body parts statistically depends on the zone, so this part of the project hypothesis is proved. It can be explained with the fact that lower limbs (legs, feet etc.) are usually being used binary as an instrument. There are no a lot of variants of singular usage of a leg or a foot. That we can not apply to the upper limbs (hands, arms, shoulders etc.), which can be a singual instrument as well as a plural instrument. 

However, the part of the hypothesis of depending on a verb cluster doesn't seem to be true. At least according to the data I have. The idea was that a predicate determines an argument and some of the constructions had clear boundary conditions (f.e. when it's possible to use only plural or only singular noun). However, there is not enough data to find any clear statistical dependencies.
The cluster "IDIOM" is a special one. It was the only cluster that had no possibilities to use both numbers. It can be explained with the fact that the Semantics of this constructions is related to a methaphorical meaning that is almost in all the cases morphologically fixed.


##  7. References.

Andersen, E. 1978. Lexical universals of body-part terminology. In Universals of Human Language, J. Greenberg (Ed.), 335– 368. Stanford CAC Stanford University Press.  
Bonch-Osmolovskaja, A. A., Merdanova, S. R., Rahilina, E. V., Reznikova. T. I. 2008. “Chastnye tipologii” v leksike: leksicheskoe pole boli. M.  
Bricyn, V. M., Rahilina, E. V., Reznikova, T. I., Javorskaja, G.M. 2009. Koncept bol’ v tipologicheskom osveshhenii. K.: Vidavnichiĭ dіm Dmitra Burago.
Chappell, H. & McGregor, W. 1996. (Eds). The Grammar of Inalienability. A Typological Perspective on Body Part Terms and the Part-whole Relation. Berlin: Mouton de Gruyter.  
Enfield, N.J. & Wierzbick,a A. 2002 (Eds). Body in description of emotion. Pragmatics and cognition 10(1–2) (special issue).  
Enfield, N.J. 2002. Semantic analysis of body parts in emotion terminology: Avoiding the exoticisms of ‘obstinate monosemy’ and ‘online extension’. Pragmatics and Cognition 10(1–2): 81–102.  
Enfield, N.J. 2006. Elicitation guide on parts of the body.  
Enfield, N.J., Majid, A. & van Staden, M. 2006. Cross-linguistic categorisation of the body: Introduction. In Majid et al. (Eds), 134–147.  
Fillmore, Ch. J. 1988. The mechanisms of ‘Construction Grammar’ // Proceedings of the Annual Meeting of the Berkeley  
Fillmore, Ch. J. 1989. Grammatical Construction Theory and the familiar dichotomies // R. Dietrich, C. F. Graumann (eds.) Language Processing in Social Context. Amsterdam: North-Holland/Elsevier, 17—38.  
Iomdin L. L. 2015. Konstrukcii mikrosintaksisa, obrazovannye russkoj leksemoj raz), SLAVIA, časopis pro slovanskou filologii, ročník 84, sešit 3, s. 291-306.  
Iordanskaja, L. & Paperno, S. 1996. A Russian-English Collocational Dictionary of the Human Body. Columbus OH: Slavica Publishers. (Also http://LexiconBrige.Com/for the electronic edition).  
Majid, A., Enfield, N.J., & van Staden, M. 2006. (Eds). Parts of the body: Cross-linguistic categorization. Language Sciences 28(2–3): 137–360 (special issue).  
Pavlenko, G. 2002. Emotions and the body in Russian and English. In Enfield & Wierzbicka (Eds), 207–241.  
Rahilina, E. V. 2010. Lingvistika konstrukcij. Azbukovnik, Moskva.  
Rahilina, E. V., Reznikova, T. I. 2013. Frejmovyj podhod k leksicheskoj tipologii. Voprosy jazykoznanija, No2.  


